DATA 624 Homework 9

### 2020-04-20 DATA624/HW9.Rmd
More work on HW 9

```{r}
rf_model_1 <- randomForest(y ~ ., data = simulated, importance = TRUE, ntree = 1000)
rf_imp_1 <- varImp(rf_model_1, scale = FALSE)
rf_imp_1  %>% 
  as.data.frame() %>% 
  rownames_to_column() %>% 
  rename(Variable = rowname) %>%
  arrange(desc(Overall)) %>%
  kable() %>%
  kable_styling()
```

@@

```{r}
rf_model_2 <- randomForest(y ~ ., data = simulated, importance = TRUE, ntree = 1000)
rf_imp_2 <- varImp(rf_model_2, scale = FALSE)
rf_imp_2 %>% 
  as.data.frame() %>% 
  rownames_to_column() %>% 
  rename(Variable = rowname) %>%
  arrange(desc(Overall)) %>%
  kable() %>%
  kable_styling()
```

@@

```{r}
rf_model_3 <- cforest(y ~ ., data = simulated)
rf_imp_3 <- varImp(rf_model_3, conditional = TRUE)
rf_imp_3 %>% 
  as.data.frame() %>% 
  rownames_to_column() %>% 
  rename(Variable = rowname) %>%
  arrange(desc(Overall)) %>%
  kable() %>%
  kable_styling()
```

@@

```{r}
gbm_grid <- expand.grid(.interaction.depth = seq(1, 7, by = 2),
                        .n.trees = seq(100, 1000, by = 100),
                        .shrinkage = c(0.01, 0.1),
                        .n.minobsinnode = 10)
gbm_model_1 <- train(y ~ ., data = select(simulated, -duplicate1), method="gbm", tuneGrid = gbm_grid, verbose = FALSE)

@@

```{r}
cubist_model_1 <- train(y ~ ., data = select(simulated, -duplicate1), method="cubist")

@@

***I will create a simulated dataset generated by a non-linear function.  I will then train the tree based models with varying levels of pruning.  I will then look at the MSE on the training set and test set in relation to the complexity of the tree.**

```{r}
set.seed(42)

@@

In stocastic gradient boosting the bagging fraction and learning rate will govern the construction of the trees as they are guided by the gradient.  Although the optimal values of these parameters should be obtained through the tuning process, it is helpful to understand how the magnitudes of these parameters affect magnitudes of variable importance.  Figure 8.24 provides the variable importance plots for boosting using two extreme values for the baggin fraction (0.1. and 0.9) and the learning rate (0.1 and 0.9) for the solubility data.  The left-hand plot has both parameters set to 0.1, and the right-hand plot has both set to 0.9:

@@

@@

@@

@@

@@

>>>>

```{r 8_1_A}
rf_model_1 <- randomForest(y ~ ., data = simulated, importance = TRUE, ntree = 1000)
rf_imp_1 <- varImp(rf_model_1, scale = FALSE)
rf_imp_1  %>% 
  as.data.frame() %>% 
  rownames_to_column() %>% 
  rename(Variable = rowname) %>%
  arrange(desc(Overall)) %>%
  kable() %>%
  kable_styling()
```

@@

```{r 8_1_B}
rf_model_2 <- randomForest(y ~ ., data = simulated, importance = TRUE, ntree = 1000)
rf_imp_2 <- varImp(rf_model_2, scale = FALSE)
rf_imp_2 %>% 
  as.data.frame() %>% 
  rownames_to_column() %>% 
  rename(Variable = rowname) %>%
  arrange(desc(Overall)) %>%
  kable() %>%
  kable_styling()
```

@@

```{r 8_1_C}
rf_model_3 <- cforest(y ~ ., data = simulated)
rf_imp_3 <- varImp(rf_model_3, conditional = TRUE)
rf_imp_3 %>% 
  as.data.frame() %>% 
  rownames_to_column() %>% 
  rename(Variable = rowname) %>%
  arrange(desc(Overall)) %>%
  kable() %>%
  kable_styling()
```

@@

```{r 8_1_D_gbm}
gbm_grid <- expand.grid(.interaction.depth = seq(1, 7, by = 2),
                        .n.trees = seq(100, 1000, by = 100),
                        .shrinkage = c(0.01, 0.1),
                        .n.minobsinnode = 10)
gbm_model_1 <- train(y ~ ., data = select(simulated, -duplicate1), method="gbm", tuneGrid = gbm_grid, verbose = FALSE)

@@

```{r 8_1_D_cubist}
cubist_model_1 <- train(y ~ ., data = select(simulated, -duplicate1), method="cubist")

@@

**I will create a simulated dataset generated by a non-linear function.  I will then train the tree based models with varying levels of pruning.  I will then look at the MSE on the training set and test set in relation to the complexity of the tree.**

```{r 8_2}
set.seed(42)

@@

In stocastic gradient boosting the bagging fraction and learning rate will govern the construction of the trees as they are guided by the gradient.  Although the optimal values of these parameters should be obtained through the tuning process, it is helpful to understand how the magnitudes of these parameters affect magnitudes of variable importance.  Figure 8.24 provides the variable importance plots for boosting using two extreme values for the bagging fraction (0.1. and 0.9) and the learning rate (0.1 and 0.9) for the solubility data.  The left-hand plot has both parameters set to 0.1, and the right-hand plot has both set to 0.9:

![Fig 8.24](fig_8_24.png)

@@

**Because the learning rate is set to 0.1, the importance get's spread out over more predictors.  The higher learning rate will focus the importance on a smaller set of variables.**

@@

**The one on the left.  It will generalize while the one on the right will overfit the training data.  Always go for an ensemble of weak predictors.**

@@

**The predictor importance would get spread across more predictors as the interaction depth would increase.  The slope would decrease.**

@@

```{r}
library(AppliedPredictiveModeling)
data(ChemicalManufacturingProcess)

# Make this reproducible
set.seed(42)

knn_model <- preProcess(ChemicalManufacturingProcess, "knnImpute")
df <- predict(knn_model, ChemicalManufacturingProcess)

df <- df %>%
  select_at(vars(-one_of(nearZeroVar(., names = TRUE))))

in_train <- createDataPartition(df$Yield, times = 1, p = 0.8, list = FALSE)
train_df <- df[in_train, ]
test_df <- df[-in_train, ]

pls_model <- train(
  Yield ~ ., data = train_df, method = "pls",
  center = TRUE,
  scale = TRUE,
  trControl = trainControl("cv", number = 10),
  tuneLength = 25
)

pls_predictions <- predict(pls_model, test_df)

pls_in_sample <- pls_model$results[pls_model$results$ncomp == pls_model$bestTune$ncomp,]
results <- data.frame(t(postResample(pred = pls_predictions, obs = test_df$Yield))) %>%
  mutate("In Sample RMSE" = pls_in_sample$RMSE,
         "In Sample Rsquared" = pls_in_sample$Rsquared,
         "In Sample MAE" = pls_in_sample$MAE,
         "Model"= "PLS")
```

**I'm going to fit the data against a series of tree based models.  I will use caret with vanilla settings to see how well the preform against a tuned PLS model.**

```{r}
gbm_model <- train(Yield ~ ., data = train_df, method="gbm", verbose = FALSE)
gbm_predictions <- predict(gbm_model, test_df)
gbm_in_sample <- merge(gbm_model$results, gbm_model$bestTune)
results <- data.frame(t(postResample(pred = gbm_predictions, obs = test_df$Yield))) %>%
  mutate("In Sample RMSE" = gbm_in_sample$RMSE,
         "In Sample Rsquared" = gbm_in_sample$Rsquared,
         "In Sample MAE" = gbm_in_sample$MAE,
         "Model"= "Boosted Tree") %>%
  rbind(results)
```

#### Random Forest

```{r}
rf_model <- train(Yield ~ ., data = train_df, method = "rf")
rf_predictions <- predict(rf_model, test_df)
rf_in_sample <- merge(rf_model$results, rf_model$bestTune)
results <- data.frame(t(postResample(pred = rf_predictions, obs = test_df$Yield))) %>%
  mutate("In Sample RMSE" = rf_in_sample$RMSE,
         "In Sample Rsquared" = rf_in_sample$Rsquared,
         "In Sample MAE" = rf_in_sample$MAE,
         "Model"= "Random Forest") %>%
  rbind(results)
```

#### Cubist Model

```{r}
cubist_model <- train(Yield ~ ., data = train_df, method="cubist")

cubist_predictions <- predict(cubist_model, test_df)
cubist_in_sample <- merge(cubist_model$results, cubist_model$bestTune)
results <- data.frame(t(postResample(pred = cubist_predictions, obs = test_df$Yield))) %>%
  mutate("In Sample RMSE" = cubist_in_sample$RMSE,
         "In Sample Rsquared" = cubist_in_sample$Rsquared,
         "In Sample MAE" = cubist_in_sample$MAE,
         "Model"= "Cubist") %>%
  rbind(results)
```

#### Results

```{r}
results %>%
  kable() %>%
  kable_styling()
```

**All tree model's fit the in sample data better than the PLS model.  However they may have overfit the data as the RMSE on the test set is higher for the random forest and boosted tree.  The cubist model is the only on that out preformed the PLS on the test set.  These findings are on vanilla models.  Additional tuning could be used to improve the model's performance**

@@

**Here's the top ten predictors:**

```{r}
varImp(cubist_model, 10)
```

**The manufacturing process variables continue to dominate the list.  Some of the variables on on both lists (in different orders of importance) while there are differences between the two lists:**

```{r}
varImp(pls_model, 10)
```
